---
title: "NYCRace"
author: "Cinny Lin"
date: "7/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Plot New York City race distribution with [tidycensus](https://walker-data.com/tidycensus/articles/basic-usage.html)

## install packages
- tidycensus 
You need to set a census API key, by entering census_api_key("YOUR API KEY GOES HERE") into the console. To request a census API key, go here: https://api.census.gov/data/key_signup.html.

## import libraries
```{r libraries}
library(tidyverse)
library(janitor)
library(tidycensus)
```

## Getting Datasets

There are two major functions implemented in **tidycensus**: 
- `get_decennial` grants access to the `1990`, `2000`, and `2010` decennial US Census APIs
- `get_acs` grants access to the 5-year American Community Survey APIs

To view/search for variables, run `load_variables(YEAR, DATASET, cache = TRUE)`.
- `DATASET` is one of "sf1", "sf3", or "acs5".
- Assign the result to a variable and set `cache = TRUE` to store the result on your computer for future access, and use the View function in RStudio to interactively browse for variables.

Get race codes.

```{r race}
racevars <- c(White = "B02001_002", 
              Black = "B02001_003", 
              Asian = "B02001_005",
              Hispanic = "B03003_003")
```

`get_acs()` allows us to grab American Community Survey (ACS) data.
`tract` is a statistical quantity. 

```{r acs data}
harris <- get_acs(geography = "tract",
                  variables = racevars, 
                  year = 2009,
                  state = "NY",
                  county = "New York County", #Manhattan
                  geometry = TRUE,
                  summary_var = "B02001_001")
```

Wrangling and mutating data above.

### syntax
- `facet_wrap` means creating different plots side-by-side separated by variable. 
In this case it will create four plots based on the race values we took from the ACS data.

- `geom_sf()` is for mapping functionality.

- `jpeg("plot.jpg")` to autosave file and `dev.off()` to close file.

```{r map}
# open jpg file
jpeg(filename="NYCRace_2014_2018.jpg", width=600, height=800)

# create plot
harris %>%
  # wrangle data
  mutate(Percent = 100 * (estimate / summary_est)) %>%
  
  # plot
  ggplot(aes(fill = Percent, color = Percent)) +
  facet_wrap(~ variable) +
  geom_sf() +
  
  # annotate/customize
  scale_fill_viridis_c(direction = -1) +
  scale_color_viridis_c(direction = -1) +
  labs(title = "Racial Geography of Manhattan, NY",
       caption = "Source: American Community Survey 2014-2018") +
       #caption = "Source: American Community Survey 2009-2013") +
  theme_void()


# close file
dev.off()
```




























